---
title: "Exploring"
output: html_document
date: "2024-02-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(plotly)
library(urbnmapr)
options(scipen = 99)
```

```{r}
install.packages("devtools")
devtools::install_github("UrbanInstitute/urbnmapr")
```


*Original dataframe from N-MHSS Codebook*
```{r}
nmhss_2019 <- read_csv(here("nmhss-puf-2019-csv.csv"))
```


*Omitted nonessential variables from original dataframe, created nmhss_2019_2*
```{r}
nmhss_2019_2 <- nmhss_2019[, !(names(nmhss_2019) %in% c('CASEID','LANGPROV','LANG16','LANG_B','LANG1','LANG2','LANG3','LANG21','LANG4','LANG5','LANG6','LANG7','LANG8','LANG24','LANG9','LANG10','LANG22','LANG25','LANG26','LANG11','LANG19','LANG23','LANG12','LANG13','LANG14','LANG15','LANG20','LANG17','LANG18','SMOKINGPOLICY'))]

```


*Created a new dataframe to include data from HI only*
```{r}
row_numbers_to_include <- 3100:3138
nmhss_2019_HI <- nmhss_2019_2[row_numbers_to_include, ]

```


*Isolated for primary treatment focus on 'Mental health treatment' (value of 1) and 'Mix of mental health and substance use treatment' (value of 3) facilities - Question A8 in N-MHSS Survey*
```{r}
mhsu_facilities_2019_HI <- subset(nmhss_2019_HI, FOCUS != 4) 

```





*Separated data by ownership, as defined by the N-MHSS Codebook, these are "Organization that operates the facility". Parameters are: 1 - Private for-profit organization; 2 - Private non-profit organization; 3 - Public agency or department - Question A10 in N-MHSS Survey. Omitted 'PUBLICAGENCY' variable as this is a subset variable for Q.A10 (Q.A10a)*
```{r}
private_for_profit <- mhsu_facilities_2019_HI[mhsu_facilities_2019_HI$OWNERSHP == 1,!(names(mhsu_facilities_2019_HI) %in% c('PUBLICAGENCY', 'FOCUS','OWNERSHP')) ]

```

```{r}
private_non_profit <- mhsu_facilities_2019_HI[mhsu_facilities_2019_HI$OWNERSHP == 2,!(names(mhsu_facilities_2019_HI) %in% c('PUBLICAGENCY','FOCUS','OWNERSHP')) ]

```

```{r}
public_agency_dept <- mhsu_facilities_2019_HI[mhsu_facilities_2019_HI$OWNERSHP == 3,!(names(mhsu_facilities_2019_HI) %in% c('FOCUS','OWNERSHP','PUBLICAGENCY')) ]

```


*Facilities by organization* 
```{r}
ggplot(data=mhsu_facilities_2019_HI, aes(x=OWNERSHP))+
  geom_bar(stat="count")
```
```{r}
mhsu_facilities_2019_HI$OWNERSHP[mhsu_facilities_2019_HI$OWNERSHP == "1"] <- "Private, for profit organization"
mhsu_facilities_2019_HI$OWNERSHP[mhsu_facilities_2019_HI$OWNERSHP == "2"] <- "Private, non profit organization"
mhsu_facilities_2019_HI$OWNERSHP[mhsu_facilities_2019_HI$OWNERSHP == "3"] <- "Public agency organization"
```



```{r}
mhsu_facilities_2019_HI$PUBLICAGENCY <- as.factor(mhsu_facilities_2019_HI$PUBLICAGENCY)
```




```{r}
ggplot(data=mhsu_facilities_2019_HI, aes(x=OWNERSHP, fill=PUBLICAGENCY))+
  geom_bar(stat="count")+
  theme_minimal()+
  labs(title= "Facilities by Organizations in Hawaii",
       x="Ownership",
       y="Number of facilities")
```
```{r}
nmhss_2019_2$PUBLICAGENCY <- as.factor(nmhss_2019_2$PUBLICAGENCY)
```


```{r}
facility_organization_plot <- ggplot(data=nmhss_2019_2, aes(x=OWNERSHP, fill=PUBLICAGENCY))+
  geom_bar(stat="count")+
  theme_minimal()+
  labs(title= "Facilities by Organizations",
       x="Ownership",
       y="Number of facilities")
ggplotly(facility_organization_plot)
```


```{r}
ggplot(data=nmhss_2019_2, aes(x=OWNERSHP, fill=PUBLICAGENCY))+
  geom_bar(stat="count")+
  theme_minimal()+
  labs(title= "Facilities by Organizations",
       x="Ownership",
       y="Number of facilities")+
  facet_wrap(~LST)
```



*Facility Type by Ownership*

```{r}
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "1"] <- "Psychiatric hospital"
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "2"] <- "Separate inpatient psychiatric unit of a general hospital"
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "3"] <- "Residential treatment center for children"
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "4"] <- "Residential treatment center for adults"
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "5"] <- "Other type of residential treatment facility"
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "6"] <- "Veterans Administration Medical Center"
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "7"] <- "Community Mental Health Center"
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "8"] <- "Certified Community Behavioral Health Clinic"
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "9"] <- "Partial hospitalization/day treatment facility"
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "10"] <- "Outpatient mental health facility"
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "11"] <- "Multi-setting mental health facility"
mhsu_facilities_2019_HI$FACILITYTYPE[mhsu_facilities_2019_HI$FACILITYTYPE == "12"] <- "Other"
```

```{r}
mhsu_facilities_2019_HI$FACILITYTYPE <- as.factor(mhsu_facilities_2019_HI$FACILITYTYPE)
```

```{r}
ggplot(data=mhsu_facilities_2019_HI, aes(x=OWNERSHP, fill=FACILITYTYPE))+
  geom_bar(stat="count")+
  theme_minimal()+
  labs(title= "Facility Type by Ownership",
       x="Ownership",
       y="Number of facilities")
```


```{r}
nmhss_2019_2$FACILITYTYPE <- as.factor(nmhss_2019_2$FACILITYTYPE)
```

```{r}
facility_organization_plot <- ggplot(data=nmhss_2019_2, aes(x=OWNERSHP, fill=FACILITYTYPE))+
  geom_bar(stat="count")+
  theme_minimal()+
  labs(title= "Facility Type by Ownership",
       x="Ownership",
       y="Number of facilities")
ggplotly(facility_organization_plot)
```


```{r}
facilitytype_by_ownership <- ggplot(data=nmhss_2019_2, aes(x=OWNERSHP, fill=FACILITYTYPE))+
  geom_bar(stat="count")+
  theme_minimal()+
  labs(title= "Facility Type by Ownership",
       x="Ownership",
       y="Number of facilities")+
  facet_wrap(~LST)
ggplotly(facilitytype_by_ownership)
```

*Adding population data for each state*


```{r}
us_state_population_2019 <- read_csv(here("us-states-territories.csv"))

```
*Joining population data with nmhss data*


```{r}
colnames(nmhss_2019_2)[colnames(nmhss_2019_2) == "LST"] ="Abbreviation"
```

```{r}
nmhss_pop <- full_join(nmhss_2019_2, us_state_population_2019, by="Abbreviation")
```

```{r}
colnames(nmhss_pop)[colnames(nmhss_pop) == "Population (2019)"] ="pop_2019"
```



*number of facilities by state/population*




```{r}
nmhss_pop_counts <- nmhss_pop %>% 
  group_by(Abbreviation, pop_2019) %>% 
  count()
```



```{r}
pop_facility_counts <- ggplot(nmhss_pop_counts, aes(x = n, pop_2019, label = Abbreviation)) +
  geom_point() +
  theme_minimal()

ggplotly(pop_facility_counts)
  
  
```

```{r}
nmhss_pop_counts_per_capita <- nmhss_pop_counts %>% 
  group_by(Abbreviation, pop_2019) %>% 
  summarise(per_capita = n/pop_2019)
```


```{r}
ggplot(nmhss_pop_counts_per_capita, aes(y = reorder(Abbreviation, per_capita),
                                        x = per_capita)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "facilities per capita",
       y = "") +
  theme_minimal()
```



Challenge 



Can you use urbnmapr to make a map this information?


```{r}
map_data <- get_urbn_map(map = "states", sf = TRUE)
```

change the column name state_abbv in map_data to Abbreviation like the way it is in nmhss_pop_counts_per_capita

```{r}
colnames(map_data)[colnames(map_data) == "state_abbv"] ="Abbreviation"
```


then you should be able to join the data based on the shared column name (Abbreviation)

```{r}
nmhss_map <- full_join(map_data, nmhss_pop_counts_per_capita, by="Abbreviation")
```


then you should be able to plot!

```{r}

ggplot() +
  geom_sf(nmhss_map,
          mapping = aes(fill = per_capita),
          color = "white", size = 0.25) +
  labs(fill = "Facilities per Capita") + 
  scale_fill_viridis_c(option="magma") + 
  theme_minimal()
```



`

