---
title: "Exploring"
output: html_document
date: "2024-02-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
```


*Original dataframe from N-MHSS Codebook*
```{r}
nmhss_2019 <- read_csv(here("nmhss-puf-2019-csv.csv"))
```


*Omitted nonessential variables from original dataframe, created nmhss_2019_2*
```{r}
nmhss_2019_2 <- nmhss_2019[, !(names(nmhss_2019) %in% c('CASEID','LANGPROV','LANG16','LANG_B','LANG1','LANG2','LANG3','LANG21','LANG4','LANG5','LANG6','LANG7','LANG8','LANG24','LANG9','LANG10','LANG22','LANG25','LANG26','LANG11','LANG19','LANG23','LANG12','LANG13','LANG14','LANG15','LANG20','LANG17','LANG18','SMOKINGPOLICY'))]

```


*Created a new dataframe to include data from HI only*
```{r}
row_numbers_to_include <- 3100:3138
nmhss_2019_HI <- nmhss_2019_2[row_numbers_to_include, ]

```


*Isolated for primary treatment focus on 'Mental health treatment' (value of 1) and 'Mix of mental health and substance use treatment' (value of 3) facilities - Question A8 in N-MHSS Survey*
```{r}
mhsu_facilities_2019_HI <- subset(nmhss_2019_HI, FOCUS != 4) 

```





*Separated data by ownership, as defined by the N-MHSS Codebook, these are "Organization that operates the facility". Parameters are: 1 - Private for-profit organization; 2 - Private non-profit organization; 3 - Public agency or department - Question A10 in N-MHSS Survey. Omitted 'PUBLICAGENCY' variable as this is a subset variable for Q.A10 (Q.A10a)*
```{r}
private_for_profit <- mhsu_facilities_2019_HI[mhsu_facilities_2019_HI$OWNERSHP == 1,!(names(mhsu_facilities_2019_HI) %in% c('PUBLICAGENCY', 'FOCUS','OWNERSHP')) ]

```

```{r}
private_non_profit <- mhsu_facilities_2019_HI[mhsu_facilities_2019_HI$OWNERSHP == 2,!(names(mhsu_facilities_2019_HI) %in% c('PUBLICAGENCY','FOCUS','OWNERSHP')) ]

```

```{r}
public_agency_dept <- mhsu_facilities_2019_HI[mhsu_facilities_2019_HI$OWNERSHP == 3,!(names(mhsu_facilities_2019_HI) %in% c('FOCUS','OWNERSHP','PUBLICAGENCY')) ]

```














